Learning Action Models from State Observations

### 2018-01-30 doc/ijcai-18/observations-learning.tex
Minor changes

Motivated by recent advances on the synthesis of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, we present a novel planning compilation for learning \strips\ action models from state observations. A solution to the classical planning task output by our compilation is a sequence of actions that determines the preconditions/effects of the \strips\ operator schemas (the learned action model). \textbf{Learning action models as planning} leverages off-the-shelf planners and opens up a way towards the \emph{bootstrapping} of planning action models, enabling a planner to gradually learn and update its action model. The practicality of the compilation approach allow us to report learning results over fifteen IPC planning domains.

For the training samples, we adopt a middle ground between unstructured inputs and plan traces, wherein only state observations are required, without further information about the state transitions. \textbf{Learning from state observations} is a relevant advancement as in many applications the actual actions executed by the observed agent are not available but instead, we can observe the resulting states. Learning action models from state observations broadens then the range of application to external observers and facilitates the representation of imperfect observability (as shown in plan recognition \cite{SohrabiRU16}) as well as the learning from unstructured data (like state images \cite{AsaiF18}).

@@

\begin{lemma}
Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves the $\Lambda$ learning task.
\end{lemma}

@@

The size of the classcial planning task $P_{\Lambda}$ output by the compilation depends on:
\begin{itemize}
\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ given as input of the $\Lambda$ learning task. The larger these numbers, the larget the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding actions for {\em programming} operator schema $\xi\in\mathcal{M}$.
\item The number of given state observations, that is $|\mathcal{O}|$. The larger this number, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
\end{itemize}

@@

To illustrate this, the plan of Figure~\ref{fig:plan} solves the task of editing a \emph{blocksworld} action model $\mathcal{M}$ where the positive effects {\tt(handempty)} and {\tt (clear ?v1)} of the {\tt\small stack} schema are missing. This plan edits the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited model $\mathcal{M}'$ in the five-observation sequence of Figure~\ref{fig:lexample}.

>>>>

Motivated by recent advances on the synthesis of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, we present a novel planning compilation for learning \strips\ action models from state observations. A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions/effects of the \strips\ operator schemas (the learned action model). \textbf{Learning action models as planning} leverages off-the-shelf planners and opens up a way towards the \emph{bootstrapping} of planning action models, enabling a planner to gradually learn and update its action model. The practicality of the compilation approach allow us to report learning results over fifteen IPC planning domains.

For the training samples, we adopt a middle ground between unstructured inputs and plan traces, wherein only state observations are required, without further information about the state transitions. \textbf{Learning from state observations} is a relevant advancement as in many applications the actual actions executed by the observed agent are not available but instead, we can observe the resulting states. Learning action models from state observations broadens then the range of application to external observers and facilitates the representation of imperfect observability (as shown in plan recognition \cite{SohrabiRU16}) as well as learning from unstructured data (like state images \cite{AsaiF18}).

@@

\begin{lemma}
Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$.
\end{lemma}

@@

The size of the classical planning task $P_{\Lambda}$ output by the compilation depends on:
\begin{itemize}
\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ given as input to the $\Lambda$ learning task. The larger these numbers, the larger the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding actions for {\em programming} operator schema $\xi\in\mathcal{M}$.
\item The number of given state observations, that is $|\mathcal{O}|$. The larger this number, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
\end{itemize}

@@

To illustrate this, the plan of Figure~\ref{fig:plan} solves the task of editing a \emph{blocksworld} action model $\mathcal{M}$ where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing. This plan edits the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited model $\mathcal{M}'$ in the five-observation sequence of Figure~\ref{fig:lexample}.
